US General Society Survey Analysis

### 2024-03-09 paper/paper.qmd
included model and echo code

---
title: "US General Society Survey Analysis"
author: "Shreya Sakura Noskor"
thanks: "Code and data are available at: [repository](https://github.com/NotSakura/USSocialSurvey.git)"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
```

@@

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

@@

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

@@

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

---
title: "US General Society Survey Analysis"
thanks: "Code and data are available at: [repository](https://github.com/NotSakura/USSocialSurvey.git)"
classoption: abstract
author: "Cristina Burca, Yan Mezhiborsky, Sakura Noskor"
format: pdf
date: today
date-format: long
abstract: |
 Some abstact
include-in-header: 
  text:
    \renewcommand{\abstractname}{Abstract}
bibliography: references.bib
margins:
  top: 0.25in
  bottom: 0.25in
  left: 0.5in
  right: 0.5in
toc: true
---

@@

library(tidyverse)
```

library(dataverse)
library(tidyverse)

ces2020 <-
  get_dataframe_by_name(
    filename = "CES20_Common_OUTPUT_vv.csv",
    dataset = "10.7910/DVN/E9N6PH",
    server = "dataverse.harvard.edu",
    .f = read_csv
  ) |>
  select(votereg, CC20_410, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)

write_csv(ces2020, "ces2020.csv")

```

ces2020 <-
  read_csv(
    "ces2020.csv",
    col_types =
      cols(
        "votereg" = col_integer(),
        "CC20_410" = col_integer(),
        "employ" = col_integer(),
        "dualcit" = col_integer(), 
        "gender" = col_integer(), 
        "birthyr" = col_integer(), 
        "race" = col_integer(), 
        "inputstate" = col_integer(), 
        "presvote16post" = col_integer(), 
        "religpew" = col_integer(), 
        "edloan" = col_integer()
      )
  )

```

ces2020 <-
  ces2020 |>
  filter(votereg == 1,
         CC20_410 %in% c(1, 2)) |>
  mutate(
    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
    voted_for = as_factor(voted_for),
    gender = if_else(gender == 1, "Male", "Female"),
    employ = case_when(
      employ == 1 ~ "Full-time",
      employ == 2 ~ "Part-time",
      employ == 3 ~ "Laid off",
      employ == 4 ~ "Unemployed",
      employ == 5 ~ "Retired",
      employ == 6 ~ "Permannetly disabled", 
      employ == 7 ~ "Homemaker", 
      employ == 8 ~ "Student", 
      employ == 9 ~ "Other"
    ),
    employ = factor(
      employ,
      levels = c(
        "Full-time",
        "Part-time",
        "laid off",
        "unemployed",
        "retired",
        "permannetly disabled", 
        "Homemaker", 
        "Student", "Other"
      )
    ), 
    immstat = case_when(
      immstat == 1 ~ "immigrant and citizen",
      immstat == 2 ~ "immigrant not citizen",
      immstat == 3 ~ "born in US, but parent(s) immigrant",
      immstat == 4 ~ "parent and I born in US but grandparent(s) immigrant",
      immstat == 5 ~ "all born in US",
    ),
    immstat = factor(immstat, levels = c( "immigrant and citizen","immigrant not citizen", "born in US, but parent(s) immigrant",     "parent and I born in US but grandparent(s) immigrant", "all born in US")
    ),
    dualcit = if_else(dualcit == 1, "Yes", "No"),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern", 
      race == 7 ~ "Two or more races", 
      race == 8 ~ "Other"
    ),
    race = factor(race, levels = c(
        "White","Black","Hispanic", "Asian", "Native American", "Middle Eastern",  "Two or more races", "Other")
    ), 
    presvote16post = case_when(
      presvote16post == 1 ~ "Hillary Clinton",
      presvote16post == 2 ~ "Donald Trump",
      presvote16post == 3 ~ "Gary Johnson",
      presvote16post == 4 ~ "Jill Stein",
      presvote16post == 5 ~ "Evan McMullin",
      presvote16post == 6 ~ "Other", 
      presvote16post == 7 ~ "Did not vote"
    ),
    presvote16post = factor(presvote16post, levels = c(
        "Hillary Clinton","Donald Trump","Gary Johnson", "Jill Stein", "Evan McMullin", "Other", "Did not vote")
    ),
    religpew = case_when(
      religpew == 1 ~ "Protestant",
      religpew == 2 ~ "Roman Catholic",
      religpew == 3 ~ "Mormon",
      religpew == 4 ~ "Eastern or Greek Orthodox",
      religpew == 5 ~ "Jewish",
      religpew == 6 ~ "Muslim", 
      religpew == 7 ~ "Buddhist", 
      religpew == 8 ~ "Hindu", 
      religpew == 9 ~ "Atheist", 
      religpew == 10 ~ "Agnostic",
      religpew == 11 ~ "Nothing in Particular",
      religpew == 12 ~ "Something else"
    ),
    religpew = factor(
      religpew,
      levels = c(
        "Protestant",
        "Roman Catholic",
        "Mormon",
        "Eastern or Greek Orthodox",
        "Jewish",
        "Muslim", 
        "Buddhist", 
        "Hindu", "Atheist", "Agnostic", "Nothing in Particular", "Something else"
      )
    ), 
    edloan = case_when(
      edloan == 1 ~ "Yes",
      edloan == 2 ~ "No",
      edloan == 3 ~ "Other",
      edloan == 4 ~ "Don't know",
      edloan == 5 ~ "None of these",
    ),
    edloan = factor(edloan, levels = c( "Yes","No", "Other", "Don't Know", "None of these")
    ) #########ADD input state it if you want to (sakura)
  ) |>
  select(votereg, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)

```

@@

- included variables of gender, birth year, race, registered state, employment, education loans, immigration status, dual-citizenship, religion, and 2016 US vote

@@

The variable 'presvote16post' captures the 2016 US Presidential votes of individuals in the United States. Due to the minimal votes for Gary Johnson, Jill Stein, Evan McMullin, and all other candidates any votes not attributed to 'Hillary Clinton' or 'Donald Trump' were categorized as 'Other'.

We have modeled the following logisitc regression in @fig-votes:

```{r, fig.height=4, fig.width=8}
#| echo: false
#| warning: false
#| message: false
#| label: fig-votes
#| fig-cap: 'Logistic regression of 2016 US Presidential votes comparing parameters of gender and employment status'

#logisitc regression for binary outcome: YES OR NO, THIS OR THAT

#choose values "Hillary Clinton" and "Donald Trump", all other values as "Other"
ces2020_votes <- ces2020 %>%
  filter(presvote16post != 'Did not vote') %>%
  mutate(presvote16post = case_when(
    presvote16post == "Hillary Clinton" ~ "Hillary Clinton",
    presvote16post == "Donald Trump" ~ "Donald Trump",
    presvote16post == "Gary Johnson" ~ "Other",
    presvote16post == "Jill Stein" ~ "Other",
    presvote16post == "Evan McMullin" ~ "Other",
    presvote16post == "Other" ~ "Other"
  ))

#plot 2016 votes against employment status and gender
#SO MANY NA VALUES???
ces2020_votes %>%
  ggplot(aes(x= employ, fill=presvote16post)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender)) +
  theme_minimal() +
  coord_flip() +
  labs(x ="Employment status",
      y = "Count of votes",
      fill = "Vote in 2016")

```

@@

```

@@

@@

```
